---
title: "Codes: A Real Time Monitoring Approach for Bivariate Event Data"
author: "Inez Maria Zwetsloot, Tahir Mahmood, Funmilola Mary Taiwo and Zezhong Wang"
date: "22 June 2021"
output: pdf_document
bibliography: references.bib 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file contains codes to reproduce the results of our paper @Zwetsloot2021.

```{r install-packages, include=FALSE, eval=FALSE}
install.packages(knitr)
install.packages(dplyr)
install.packages(VGAM)
install.packages(tidyverse)
```

```{r load-packages, include=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
library(knitr)      ## For Writing operations
library(dplyr)      ## For Data operations
library(VGAM)       ## For Lambert W function
library(tidyverse)  ## For Writing operations
```

# 1. Required Functions

## 1.1 Function for GBE ATS when $\delta \neq 1$

### 1.1.1. Function to generate GBE random numbers
```{r data-model-GBE}
# general random data from bivariate GBE model  
rgbe = function(n, par){
  U = runif(n,min=0,max=1)
  M = rbinom(n,size=1,prob=par[3])
  V1 = rexp(n, rate=1)
  V2 = rexp(n, rate=1)
  V = V1 + M*V2
  X1 = par[1]*(U^par[3])*V
  X2 = par[2]*((1-U)^par[3])*V 
  X = matrix(data = NA,nrow=2,ncol=n)
  X[1,] = X1
  X[2,] = X2
  return(X)
}
```

### 1.1.2. Function for $C(x,y)$
```{r fC}
fC = function(x,y,par){
  t1 = par[1];t2 = par[2];del = par[3]
  C = (x/t1)^(1/del) + (y/t2)^(1/del)
  return(C)  
}
```


### 1.1.3. Function for GBE limits
```{r, CL.GBE}
CL.Z =  function(par,ATS){
  t1=par[1];t2=par[2];del=par[3]
  C11=fC(1,1,par)
  ETBE0=0.5*(t1+t2-(1/(C11^del)))
  alpha = ETBE0/ATS
  ucl= -(1/(C11^del))*log(alpha)
  return(ucl)
}

CL.M =  function(par,ATS,vX){
  t1=par[1];t2=par[2];del=par[3]
  C11=fC(1,1,par)
  ETBE0=0.5*(t1+t2-(C11^(-del)))
  alpha = ETBE0/ATS
  z = min(vX)
  v = as.numeric(vX[1]<vX[2])

  if (del==1){
    if (v==1){
       ucl = z - t2*log(alpha)
    } else if (v==0){
       ucl = z - t1*log(alpha)
    }
  } else if (del<1){
    if (v==1){
        G1 = (del/(1-del))*(fC(z,z,par)^del)*((1-alpha)*exp(-(fC(z,z,par)^del)))^(-del/(1-del))
        G2 = (del/(1-del))*(fC(z,z,par)^del)*(alpha*exp(-(fC(z,z,par)^del)))^(-del/(1-del))
        WG1 = lambertW(G1,tolerance = 1e-10,maxit=50)
        WG2 = lambertW(G2,tolerance = 1e-10,maxit=50)
        ucl = ((t2*(1-del)/del*WG2)^(1/del)-(z*t2/t1)^(1/del))^del       
    } else if (v==0){
        G1 = (del/(1-del))*(fC(z,z,par)^del)*((1-alpha)*exp(-(fC(z,z,par)^del)))^(-del/(1-del))
        G2 = (del/(1-del))*(fC(z,z,par)^del)*(alpha*exp(-(fC(z,z,par)^del)))^(-del/(1-del))
        WG1 = lambertW(G1,tolerance = 1e-10,maxit=50)
        WG2 = lambertW(G2,tolerance = 1e-10,maxit=50)
        ucl = ((t1*(1-del)/del*WG2)^(1/del)-(z*t1/t2)^(1/del))^del    
    } 
  }
  return(ucl)
}
```



### 1.1.4. Function based on simulations to compute ATS
```{r mtbe.ats}
GBE_ATS = function(ATS,par0,par1){
  R=10000;vTS=rep(0,R)
  for (i in 1:R){
    #run chart as long as there is no signal
    vSignal = c(FALSE,FALSE)
    while(all(!vSignal)){ 
     vX = rgbe(n=1,par1) #generate data from MOBW
        Z = min(vX);M = max(vX)


        #update signal vector for first passage time
       UCL.z=CL.Z(par0,ATS) #ucl for Z
       UCL.m=CL.M(par0,ATS,vX) #lcl for M
        vSignal[1] = Z > UCL.z
        vSignal[2] = M > UCL.m
        
        if(all(vSignal == FALSE)){
          vTS[i] =vTS[i]+M
      } else if(vSignal[1] == TRUE){
          vTS[i] =vTS[i]+Z
        } else if(vSignal[2] == TRUE){
          vTS[i] =vTS[i]+M
        }
    }
  }
   ATS = mean(vTS)
   return(ATS)
}
```

## 1.2. Function for GBE ATS when $\delta=1$
```{r}
GBE_ATS1=function(ATS,par0,par1){
  t1=par0[1];t2=par0[2]
  C11=fC(1,1,par0)
  ETBE0=0.5*(t1+t2-C11^(-1))
  alpha = ETBE0/ATS
  
  t1_=par1[1];t2_=par1[2]
  C11_=fC(1,1,par1)
  ETBE1=0.5*(t1_+t2_-C11_^(-1))
  
  num = 2- alpha^(C11_/C11)
  Pns1 = 1-alpha^(C11_/C11)
  P1_=1/(t1_*C11_)
  P2_=1/(t2_*C11_)
  denum = 1 - Pns1*(P1_*(1-alpha^(t2/t2_))+P2_*(1-alpha^(t1/t1_)))
  ats1=num/denum*ETBE1
return(ats1)
}
```

## 1.3. Function for MOBE ATS
```{r}
MOBE_ATS=function(ATS,par0,par1){
l1= par0[2];l2=par0[3];l12=par0[1]
l=l1+l2+l12
ETBE0 = 0.5*(l2/l^2+l2/l/(l1+l12)+l1/l^2+l1/l/(l2+l12))+l12/l^2
alpha = ETBE0/(ATS)

l_1= par1[2];l_2=par1[3];l_12=par1[1]
l_=l_1+l_2+l_12
ETBE1 = 0.5*(l_2/l_^2+l_2/l_/(l_1+l_12)+l_1/l_^2+l_1/l_/(l_2+l_12))+l_12/l_^2

Noevent = (l_1+l_2)/l_*(2-alpha^(l_/l))+l_12/l_
Psignal = alpha^(l_/l)+((l_1/l_)*alpha^((l_2+l_12)/(l2+l12))+(l_2/l_)*alpha^((l_1+l_12)/(l1+l12)))*(1-alpha^(l_/l))
ARL1 = Noevent/Psignal 
ATS1 <- ARL1*ETBE1
return(ATS1)
}
```

## 1.4. Function for MOBW ATS

```{r}
MOBW_ATS=function(ATS,par0,par1){
l1= par0[3];l2=par0[4];l12=par0[2]
l=l1+l2+l12
eta=par0[1]
ETBE0 = 0.5*gamma(1+1/eta)*((l2+l12)^(-1/eta)-(l2+l12)*l^(-1-1/eta)+(l1+l12)^(-1/eta)-(l1+l12)*l^(-1-1/eta)+2*l12*l^(-1-1/eta))
alpha=ETBE0/(2*ATS)

l_1= par1[3];l_2=par1[4];l_12=par1[2]
l_=l_1+l_2+l_12
eta_ =par1[1]
ETBE1 = 0.5*gamma(1+1/eta_)*((l_2+l_12)^(-1/eta_)-(l_2+l_12)*l_^(-1-1/eta_)+(l_1+l_12)^(-1/eta_)-(l_1+l_12)*l_^(-1-1/eta_)+2*l_12*l_^(-1-1/eta_))
L <- l_/l
L1 <- (l_1+l_12)/(l1+l12)
L2 <- (l_2+l_12)/(l2+l12)
alpha_=alpha^L+1-(1-alpha)^L

num = 1 + (l_1+l_2)/l_*(1-alpha_)
dum1 = l_1/l_*(1-(1-alpha)^L2+alpha^L2)
dum2 = l_2/l_*(1-(1-alpha)^L1+alpha^L1)
denum = alpha_ +(1-alpha_)*(dum1+dum2)
#print((1-alpha_)*(dum1+dum2))
ARL1= num/denum
ATS1 <- ARL1*ETBE1
return(ATS1)
}
```

# 2. Function to compute BTBE ATS values
```{r}
BTBE_ATS=function(ATS,par0,par1,Dist){
  if(Dist=="GBE"){
    if(par0[3]!=1){
    print("ATS is based on 10,000 Simulations")
    ATS=GBE_ATS(ATS,par0,par1)
    }else if(par1[3]==1){
    print("ATS based on analytical expression")
    ATS=GBE_ATS1(ATS,par0,par1)
    }
  }else if (Dist=="MOBE"){
    ATS=MOBE_ATS(ATS,par0,par1)
  }else if (Dist=="MOBW"){
    ATS=MOBW_ATS(ATS,par0,par1)
  }
  return(ATS)
}
```


## 2.1. Examples of BTBE chart under GBE distribution
```{r}
### When delta is equal to 1
ATS=200
par0=c(5,5,1)
par1=c(5,5,1)
BTBE_ATS(ATS,par0,par1,Dist="GBE")

### When delta is not equal to 1
ATS=200
par0=c(5,5,0.5)
par1=c(5,5,0.5)
BTBE_ATS(ATS,par0,par1,Dist="GBE")
```

## 2.2. Example of BTBE chart under MOBE distribution
```{r}
ATS=200
par0=c(1,5,5,0)
par1=c(1,5,5,0)
BTBE_ATS(ATS,par0,par1,Dist="MOBE")
```

## 2.3. Example of BTBE chart under MOBW distribution
```{r}
ATS=200
par0=c(2,5,5,0)
par1=c(2,5,5,0)
BTBE_ATS(ATS,par0,par1,Dist="MOBW")
```

# References
